---
title: "UFO"
author: "Daqi (Jen) Chen"
output: html_notebook
---

# Questions to Answer
> What is the most common shape reported in UFO sightings?
> Is there a pattern as the years go on in frequency of UFO sightings?


# Required packages 

```{r}
require(ggplot2)
require(lubridate)
require(maps)
require(mapdata)
require(data.table)

```

# Read in Data

```{r}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(maps)
library(mapdata)
library(data.table)

ufo <- read.csv("scrubbed.csv")

#####################
## Date formatting ##
#####################
ufo$datetime <- mdy_hm(ufo$datetime) 
ufo$date.posted <- mdy(ufo$date.posted)
ufo$duration..seconds.<- as.numeric(ufo$duration..seconds.)

ufo$country <- as.factor(ufo$country) 
levels(ufo$country) #[1] ""   "au" "ca" "de" "gb" "us"
levels(ufo$country) <- c("NA","Australia","Canada","Germany","Great Britain","US")
```




# Question 1:
## What is the most common shape reported in UFO sightings?

```{r}
ufo %>%
  ggplot(aes(x=forcats::fct_infreq(shape),fill=shape))+
  stat_count(show.legend = FALSE)+
  ggtitle("What is the most common shape reported in UFO sightings?")+
  theme(axis.text.x = element_text(angle=70, vjust=0.5))+
  xlab("Shape") + ylab("Frequency")
  
```

> The most common shape in UFO sightings is light.

# Question 2:
## Is there a pattern as the years go on in frequency of UFO sightings?

```{r}
ufo$year <- year(ufo$datetime)
ufo_year <- subset(ufo, year>1940)
ufo_year %>%
  ggplot(aes(x=year,y=..count..,fill=..count..))+
  stat_count(show.legend = FALSE)+
  ggtitle("Is there a pattern as the years go on in frequency of UFO sightings?")+
  xlab("Year") + ylab("Number of Sightings")
```

> As the years go on, there is a general upward trend in frequency of UFO sightings. 


